---
title: "estimate intercepts using lme4 and mplus"
author: "andrew demetriou"
---

```{r setup, include=FALSE}
library('here')            # file logistics
library('data.table')      # data frame logistics
library('tidyverse')       # code logistics
library("MplusAutomation") # Mplus integration

values <- c("POWER", "ACHIEVEMENT", "HEDONISM",  
            "STIMULATION", "SELF", "UNIVERSALISM", 
            "BENEVOLENCE", "TRADITION",  
            "CONFORMITY", "SECURITY")

# load and format datasets
response_dfs <- readRDS(file=here("VI_data_collection", "_data", "survey_response_data", "response_dfs.RDS"))
```

We have four datasets loaded into a single list:

From current survey:
*`responses_dt_c` is only complete responses from the current survey
*`responses_dt` is both all complete and all partial responses from current survey
*`responses_dt_p` is all complete responses from the pilot study
*`responses_dt_a` is all responses from current survey and pilot study

```{r}
format_df_for_mplus <- function(df){
  # select only ID columns and values scores
  mplus_df <- df %>% select(row_number, participant_ID, item_ID, all_of(values))
  
  # rename columns for Mplus
  colnames(mplus_df) <- c("row_number", "subject_ID", "item_ID", paste0("y", seq(1, 10)))
  
  # coerce to dataframe for mplus modeller
  mplus_df <- as.data.frame(mplus_df)
  
  return(mplus_df)
}
```

```{r}
fit_mplus_model <- function(df){
  
  # run mplus model
 fit <- mplusModeler(mplus_model, 
                     modelout = "VI_data_collection/_data/mplus_output/participant_model_1.0.inp", 
                     hashfilename = FALSE,
                     check = TRUE,
                     run=TRUE)

  # print errors and warnings
  print(fit$results$errors)
  print(fit$results$warnings)
  
  return(fit)
}
```

This wouldn't run in a single function, so I run it on each data set:

```{r}
## responses_dt
mplus_df <- format_df_for_mplus(response_dfs[[1]])  
  
# load mplus model specification
source(here("VI_data_collection", "3_response_intercept_analysis",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit <- mplus_df %>% fit_mplus_model()
```

```{r}
# format df
mplus_df <- format_df_for_mplus(response_dfs[[3]])  
  
# load mplus model specification
source(here("VI_data_collection", "3_response_intercept_analysis",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_p <- mplus_df %>% fit_mplus_model()
```

```{r}
# format df
mplus_df <- format_df_for_mplus(response_dfs[[2]])  
  
# load mplus model specification
source(here("VI_data_collection", "3_response_intercept_analysis",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_c <- mplus_df %>% fit_mplus_model()
```

```{r}
# format df
mplus_df <- format_df_for_mplus(response_dfs[[4]])  
  
# load mplus model specification
source(here("VI_data_collection", "3_response_intercept_analysis",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_a <- mplus_df %>% fit_mplus_model()
```
```{r}
# coerce to list
participant_fits <- list(
  fit   = fit, 
  fit_c = fit_c,
  fit_p = fit_p, 
  fit_a = fit_a
)

# save output
saveRDS(participant_fits, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_fits.RDS"))

rm(fit, fit_c, fit_p, fit_a, format_df_for_mplus, fit_mplus_model)
```

```{r}
# load mplus fit objects

# first run
#fit   <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_1.0.RDS"))
#fit_a <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_a_1.0.RDS"))
#fit_c <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_c_1.0.RDS"))
#fit_p <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_p_1.0.RDS"))

# second run
participant_fits <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_fits.RDS"))
```

```{r}
get_random_intercepts_mplus <- function(
    df, # original dataframe
    fit, # output of mplus model run
    intercept, # participant or item
    # 10 item character vector of variables in order
    DVs = c(
            "POWER", "ACHIEVEMENT", "HEDONISM",  
            "STIMULATION", "SELF", "UNIVERSALISM", 
            "BENEVOLENCE", "TRADITION",  
            "CONFORMITY", "SECURITY"
      )){
  
  intercept <- intercept |> toupper()
  
  if(intercept == 'SUBJECT_ID'){
    prefix <- 'B2a_Y'
  }else if(intercept == 'ITEM_ID'){
    prefix <- 'B2b_Y'
  }else(print("error: wrong intercept identifier"))
  
  # retrieve original IDs 
  ID_df <- df %>% dplyr::select(row_number, subject_ID, item_ID)
  
  # merge output from mplus and original IDs
  mplus_df <- fit$results$savedata
  mplus_df <- merge(mplus_df, ID_df, by.x = 'ROW_NUMB', by.y = 'row_number')
  # rename IDs to original IDs
  mplus_df$ITEM_ID <- mplus_df$item_ID
  mplus_df$SUBJECT_ID <- mplus_df$subject_ID
  
  # this is the output from the bayesian sampling process
  factor_scores <- mplus_df |>
    # .Mean represents the mean value of the resampling
    select(intercept, ends_with('.Mean'))
    
    # extract intercept SDs
    intercepts_df <- factor_scores |>
      # 'B2a' - subject intercepts 'B2b' - item intercepts
      select(intercept, starts_with(prefix))
    
    # reduce to a single row per item / participant
    #intercepts_df <- intercepts_df %>% unique()
    intercepts_df <- intercepts_df[!duplicated(intercepts_df[1]),]
    
    colnames(intercepts_df) <- c(intercept, DVs)
    
    return(intercepts_df)
}
```


```{r}
build_intercept_df <- function(df, fit_object=fit, dataset_label){
  
  # df <- responses_dt_p %>% rename(subject_ID = participant_ID)
  # fit_object <- fit_p
  
  df <- get_random_intercepts_mplus(
    # adjust column name to fit legacy function
    df %>% rename(subject_ID = participant_ID), fit_object, 
    # specify intercept to retrieve
    intercept = 'item_ID') %>%
    # pivot values columns to long format
    pivot_longer(cols = all_of(values), names_to = "value", values_to = "intercept") %>% 
    # reorder df
    arrange(value, ITEM_ID) %>%
    # add column that labels the dataset
    mutate(dataset = dataset_label)
}
```

```{r}
# build intercept df for each dataset
mplus_intercept_df_1 <- build_intercept_df(
  response_dfs[[1]], participant_fits[[1]], 
  "complete and partial responses") %>% rename(item_ID = ITEM_ID)

mplus_intercept_df_c <- build_intercept_df(
  response_dfs[[2]], participant_fits[[2]], 
  "complete responses only") %>% rename(item_ID = ITEM_ID)

mplus_intercept_df_p <- build_intercept_df(
  response_dfs[[3]], participant_fits[[3]], 
  "pilot study responses only") %>% rename(item_ID = ITEM_ID)

mplus_intercept_df_a <- build_intercept_df(
  response_dfs[[4]], participant_fits[[4]], 
  "all responses") %>% rename(item_ID = ITEM_ID)
```

```{r}
# coerce to list
mplus_item_intercept_dfs <- list(
  mplus_intercept_df_1 = mplus_intercept_df_1, 
  mplus_intercept_df_c = mplus_intercept_df_c, 
  mplus_intercept_df_p = mplus_intercept_df_p, 
  mplus_intercept_df_a = mplus_intercept_df_a)

saveRDS(mplus_item_intercept_dfs, here("VI_data_collection", "_data", "intermediary_data", "participant_mplus_item_intercept_dfs.RDS"))
```


```{r}
# load lme4 functions
source(here("VI_data_collection", "3_response_intercept_analysis", "_functions", "lme4_descriptives_function_1.7.R"))
```


```{r}
lme4_intercept_and_predicted <- function(df){
  # select only ID columns and values scores
  df <- df %>% select(participant_ID, item_ID, all_of(values))
  
  # rename columns to have same names as in functions
  colnames(df) <- c("subject_ID", "item_ID", values)
  
  # run lmer models with each value as the dependent variable and
  # extract standard deviation 95% CI for participants and songs by value
  lme4_intercept_and_predicted_dt <- df |> 
     assemble_intercept_and_predicted_dt()
}
```

```{r}
lme4_intercept_dt_1  <-lme4_intercept_and_predicted(response_dfs[[1]])   %>% 
  mutate(dataset = "complete and partial responses")
lme4_intercept_dt_c <- lme4_intercept_and_predicted(response_dfs[[2]]) %>% 
  mutate(dataset = "complete responses only")
lme4_intercept_dt_p <- lme4_intercept_and_predicted(response_dfs[[3]]) %>% 
  mutate(dataset = "pilot study responses only")
lme4_intercept_dt_a <- lme4_intercept_and_predicted(response_dfs[[4]]) %>% 
  mutate(dataset = "all responses")
```

```{r}
# coerce to list
lme4_item_intercepts_dfs <- list(
  lme4_intercept_dt_1 = lme4_intercept_dt_1, 
  lme4_intercept_dt_c = lme4_intercept_dt_c, 
  lme4_intercept_dt_p = lme4_intercept_dt_p, 
  lme4_intercept_dt_a = lme4_intercept_dt_a)

saveRDS(lme4_item_intercepts_dfs, here("VI_data_collection", "_data", "intermediary_data", "participant_lme4_item_intercept_dfs.RDS"))
```


```{r}
# quick look at correlations
df <- merge(lme4_intercept_dt, mplus_intercept_df_1, by = c("item_ID", "value"))
cor(df$intercept.x, df$intercept.y)

df <- merge(lme4_intercept_dt_c, mplus_intercept_df_c, by = c("item_ID", "value"))
cor(df$intercept.x, df$intercept.y)

df <- merge(lme4_intercept_dt_p, mplus_intercept_df_p, by = c("item_ID", "value"))
cor(df$intercept.x, df$intercept.y)

df <- merge(lme4_intercept_dt_a, mplus_intercept_df_a, by = c("item_ID", "value"))
cor(df$intercept.x, df$intercept.y)
```
