---
title: "inter rater reliability"
author: "andrew demetriou"
---

```{r setup, include=FALSE}
library('here')       # file logistics
library('data.table') # data manipulation
library('dplyr')      # data manipulation
library('tidyr')      # data manipulation
library('ggplot2')    # visualization
library('psych')      # intra class correlation
library('patchwork')  # combine plots

values <- c("POWER", "ACHIEVEMENT", "HEDONISM",  
            "STIMULATION", "SELF", "UNIVERSALISM", 
            "BENEVOLENCE", "TRADITION",  
            "CONFORMITY", "SECURITY")

#options(scipen=999)
theme_set(theme_minimal())
```

```{r}
# load datasets
response_dfs <- readRDS(file=here("VI_data_collection", "_data", "survey_response_data", "response_dfs.RDS"))
```


```{r}
# reorder columns
reorder_df <- function(df){
  df <- df %>% select(
    participant_ID, item_ID, 
    POWER, ACHIEVEMENT, HEDONISM, 
    STIMULATION, SELF, UNIVERSALISM, 
    BENEVOLENCE, TRADITION, 
    CONFORMITY, SECURITY)
}
```

```{r}
# create list of dataframes
reordered_dfs <- lapply(response_dfs, reorder_df)

rm(reorder_df)
```

```{r}
plots <- lapply(reordered_dfs, function(x){
  
  count_df <- x %>%
    group_by(item_ID) %>%
    count()

  count_median <- median(count_df$n)

  plot <- count_df %>%
    ggplot(aes(n)) +
    geom_histogram(bins=10, color = 'black', fill = 'grey') +
    xlab("ratings per item") +
    geom_vline(xintercept = count_median, color = "red")
  
  return(plot)
})

names(plots) <- names(reordered_dfs)

plots[[1]] + ggtitle(names(plots[1]))
plots[[2]] + ggtitle(names(plots[2]))

rm(plots)
```
'psych' package documentation:
"Where the data are laid out in terms of Rows (subjects) and Columns (rater or tests), the various ICCs are found by the ratio of various estimates of variance components. 
In all cases, subjects are taken as varying at random, and the residual variance is also random. The distinction between models 2 and 3 is whether the judges (items/tests) are seen as random or fixed. A further distinction is whether the emphasis is upon absolute agreement of the judges, or merely consistency."

"ICC2: A random sample of k judges rate each target. The measure is one of absolute agreement in the ratings."

Koo, Terry, and Mae Li. 2016. “A Guideline of Selecting and Reporting Intraclass Correlation Coefficients for Reliability Research.” Journal of Chiropractic Medicine 15 (March). doi:10.1016/j.jcm.2016.02.012

Definitions: inter-rater reliability: "...reflects the variation between 2 or more raters who measure the same group of subjects"

Selection criteria:
<<model selection>>
1) do we have the same set of raters for all subjects? 
No
2) do we have a sample of raters randomly selected from a larger population or a specific sample of raters?
Randomly selected from a larger population of Prolific users whose demographics represent the U.S. population. 


Two-Way Random-Effects Model:
If we randomly select our raters from a larger population of raters with similar characteristics, 2-way random-effects model is the model of choice. 
In other words, we choose 2-way random-effects model if we plan to generalize our reliability results to any raters who possess the same characteristics as the selected raters in the reliability study.


<<type selection>>
3) Are we interested in the reliability of a single rater or the mean value of multiple raters?
The mean value of multiple raters

“Type” Selection
This selection depends on how the measurement protocol will be conducted in actual application. 
For instance, if we plan to use the mean value of 3 raters as an assessment basis, the experimental design of the reliability study should involve 3 raters, and the “mean of k raters” type should be selected.

<<definition>>
4) Do we concern about consistency or agreement?
Consistency

Absolute agreement concerns if different raters assign the same score to the same subject. 
Conversely, consistency definition concerns if raters’ scores to the same group of subjects are correlated in an additive manner

```{r}
make_icc_df <- function(df){
  # iterate over list of values
  list_of_icc_dfs <- lapply(values,function(value){
  
    # select columns
  df <- df %>% dplyr::select(participant_ID, item_ID, value) %>%
    # pivot so that participants are columns
    pivot_wider(names_from = participant_ID, values_from = value)
  
  # remove item_ID column
  icc_df <- df %>% select(-item_ID) %>% 
    # pass to ICC function
    psych::ICC()
  
  # select relevant ICC output
  icc_df <- icc_df$results %>% filter(type =="ICC2k" | type == "ICC2")
  icc_df$value <- value
  return(icc_df)
})
  
  # bind list of icc dfs
  icc_df <- rbindlist(list_of_icc_dfs)
}
```

```{r}
icc_df_list <- lapply(reordered_dfs, make_icc_df)
icc_df <- rbindlist(icc_df_list, idcol = "dataset") %>%
  select(dataset, ICC, type, `lower bound`, `upper bound`, value) %>%
  rename(lower = `lower bound`, upper = `upper bound`)
#rm(make_icc_df)
```

"Based on the 95% confident interval of the ICC estimate, values less than 0.5, between 0.5 and 0.75, between 0.75 and 0.9, and greater than 0.90 are indicative of poor, moderate, good, and excellent reliability..."

```{r}
# plot two way absolute agreement vs. relative agreement by dataset
icc_df %>% 
  ggplot(aes(x = ICC, y = value, color = value)) +
  facet_wrap(~dataset) +
  geom_point() + 
  geom_errorbar(xmin = icc_df$lower, xmax =icc_df$upper) +
  theme(legend.position = "none")
```

```{r}
icc_df %>%
  filter(type == "ICC2k") %>%
  ggplot(aes(x=ICC, y = value, color = dataset, group = dataset)) +
  facet_wrap(~dataset) +
  geom_point() +
  theme(legend.position = "none") #+ 
  #geom_errorbar(xmin = icc_df$lower, xmax = icc_df$upper)
```


```{r}
responses_dt <- reordered_dfs[[1]]
```

```{r}
get_icc <- function(df=responses_dt, n, value){
  df$item_ID <- as.factor(df$item_ID)

  df%>% select(value)
  
  # get canonical mean per value
  canon <- df %>%
      #select only stimuli columns
      dplyr::select(item_ID, value, -participant_ID) %>%
      #group by each column
      group_by(item_ID) %>%
      #calculate means of all columns
      summarize_all(mean)
  
  # subsample data
  data_sample <- df %>% select(participant_ID, item_ID, value) %>% 
    group_by(item_ID) %>%
    sample_n(n) %>% ungroup() 
  
  # calculate mean sample ratings
  sample_means <- data_sample %>%
    select(-participant_ID) %>%
    group_by(item_ID) %>%
    summarize_all(mean)
  
  # calculate correlation between sample mean and canonical mean
  cor <- cor(canon[,2], sample_means[,2])[[1,1]]
  
  # compute ICC
  icc_df <- data_sample %>%
    pivot_wider(names_from = participant_ID, values_from = value) %>%
    select(-item_ID) %>% ICC() 
  
  # subset rows
  icc_df <- icc_df$results %>% filter(type == 'ICC2' | type == 'ICC2k') %>%
    select(type, ICC, `lower bound`, `upper bound`)
  
  # remove rownames
  rownames(icc_df) <- NULL
  
  # collect values to return
  icc_df$value <- value
  icc_df$n <- n
  icc_df$cor <- cor
  
  return(icc_df)
}  

#t <- get_icc(responses_dt, n=15, value=values[[2]])
```

```{r}
samples_of_ICC <- function(
  value,
  df = responses_dt, 
  n_samples = 10, 
  n_raters = seq(5, 20, by = 5)
  ){
  #draw n_raters number of samples n_samples number of times
  icc_tibble <- rep(n_raters, each = n_samples) %>% 
    purrr::map_df( function(n) { 
      get_icc(df, n, value)
    })
  return(icc_tibble)
}

#s<- samples_of_ICC(value = values[[2]])
```

```{r}
list_of_icc_dfs <- lapply(values, samples_of_ICC)
sampled_icc_df <- rbindlist(list_of_icc_dfs)
```

```{r}
# changes in ICC as a function of sample size

# ICC2
p1 <- sampled_icc_df %>%
  filter(type == "ICC2") %>%
  ggplot(aes(ICC, color = as.factor(n))) +
  facet_grid(~value)+
  #geom_density() + 
  geom_line(stat = 'density', alpha = 0.7) +
  theme(axis.text.x = element_text(angle = 90), 
        strip.text.x = element_text(size = 4)) +
  coord_flip()

# ICC2k
p2 <- sampled_icc_df %>%
  filter(type == "ICC2k") %>%
  ggplot(aes(ICC, color = as.factor(n))) +
  facet_grid(~value)+
  #geom_density() + 
  geom_line(stat = 'density', alpha = 0.7) +
  theme(axis.text.x = element_text(angle = 90), 
        strip.text.x = element_text(size = 4)) +
  coord_flip()

#p1 / p2

p2
```

```{r}
# changes in correlation to canonical mean as a function of sample size

# it actually doesn't matter which ICC is chosen, because that's not the metric of interest, but I compute both plots anyway
p3 <- sampled_icc_df %>%
  filter(type == "ICC2") %>%
  ggplot(aes(cor, color = as.factor(n))) +
  facet_grid(~value)+
  #geom_density(alpha = 0.3) + 
  geom_line(stat = 'density', alpha = .7) +
  theme(axis.text.x = element_text(angle = 90), 
        strip.text.x = element_text(size = 4)) +
  coord_flip()

p4 <- sampled_icc_df %>%
  filter(type == "ICC2k") %>%
  ggplot(aes(cor, color = as.factor(n))) +
  facet_grid(~value)+
  #geom_density(alpha = 0.3) + 
  geom_line(stat = 'density', alpha = .9) +
  theme(axis.text.x = element_text(angle = 90), 
        strip.text.x = element_text(size = 4)) +
  coord_flip() +
  labs(x = "correlation", colour = "n")


p4
```



