---
title: "machine dimension exploration"
---

```{r}
library('here')            # file logistics
library('data.table')      # data manipulation
library('tidyverse')       # data manipulation
library("psych")           # factor analysis
library("ggplot2")         # visualization
library("corrplot")        # correlation plot
library("ggridges")        # joyplots
library('patchwork')       # multiple plots 

options(scipen=999)
theme_set(theme_minimal())

# load datasets
machines_as_subjects_df <- readRDS(here("VI_data_collection", "_data", "machine_response_data", "machines_as_subjects_df.RDS"))

values <- c("POWER", "ACHIEVEMENT", "HEDONISM", 
            "STIMULATION", "SELF", "UNIVERSALISM", 
            "BENEVOLENCE", "TRADITION", 
            "CONFORMITY", "SECURITY")

# pivot to wide: each machine / value combination is a column
machines_as_fixed_df <- machines_as_subjects_df %>%
  pivot_wider(id_cols = item_ID, names_from = participant_ID, values_from = all_of(values))

# df with reduced columns for the mxm-trained models
load(file = here("VI_data_collection", "_data","intermediary_data", "machines_as_fixed_r_df.RDS"))

rm(machines_as_subjects_df)
```

```{r}
# add glove and word2vec scores to dimension reduced dataframe
machines_as_fixed_df <- merge(machines_as_fixed_r_df, machines_as_fixed_df %>% select(item_ID, contains("glove"), contains("google")), by="item_ID")

#rm(machines_as_fixed_r_df)
```

Here I examine the factors comprised of all the machine setups and the means. 

```{r}
cors <-lapply(values, function(x){
  #x<-values[2]
  
  # made df of columns that were averaged
  means_df <- machines_as_fixed_df %>% select(starts_with(x)) %>%
    select(contains("mean"))
  
  # make df of the factors
  factors_df <- machines_as_fixed_df %>% select(starts_with(x)) %>%
    select(-contains("mean"), -contains("google"), -contains("glove"))
  
  # create empty list to return
  cors <- list()
  
  # for each column in one df, compute the correlation and return it
  for(col in seq(1:4)){
    cors[col] <- cor(means_df[,col], factors_df[,col])
  }
  return(cors)
})

cors_df <- rbindlist(cors)
cors_df$value <- values
colnames(cors_df) <- c("far_idf", "far_uniform", 
                  "split_idf", "split_uniform", 
                  "values")

rm(means_df, factors_df, col, cors)
```

Here I plot a heatmap of the correlations:

```{r}
cors_df %>% pivot_longer(
  cols =  c("far_idf", "far_uniform", 
                  "split_idf", "split_uniform"),
  names_to = "model", 
  values_to = "correlation"
  ) %>%
  ggplot(., aes(x=values, y=model, fill = correlation)) +
    geom_tile(color = "white") + 
    geom_text(aes(label = round(correlation, 2)), color = "white")+
  scale_fill_gradient(low = "blue", high = "red") +
    coord_flip()

rm(cors_df)
```

High correlations for all except Hedonism. 
Here I plot the distributions:

```{r}
p1 <- machines_as_fixed_r_df %>% 
  select(item_ID, starts_with("HEDONISM")) %>%
  pivot_longer(
    cols = starts_with("HEDONISM"),
    names_to = "model", 
    values_to = "score"
  ) %>%
  group_by(model) %>%
  ggplot(aes(x= score, y = model, fill =model, alpha = 0.7)) +
  geom_density_ridges() + 
  theme(legend.position = "none") + 
  xlim(-5, 5) +
  xlab("score") + 
  ylab("model")

p2 <- machines_as_fixed_r_df %>% 
  select(item_ID, starts_with("POWER")) %>%
  pivot_longer(
    cols = starts_with("POWER"),
    names_to = "model", 
    values_to = "score"
  ) %>%
  group_by(model) %>%
  ggplot(aes(x= score, y = model, fill =model, alpha = 0.7)) +
  geom_density_ridges() + 
  theme(legend.position = "none") + 
  xlim(-5, 5) +
  xlab("score") + 
  ylab("model")

p1 / p2
```

```{r}
machines_as_fixed_df_long <- machines_as_fixed_df %>%
  pivot_longer(cols = -item_ID, values_to = "estimate", names_sep = "_", names_to = c("value", "model", "weight", "dimension_reduction"))

machines_as_fixed_df_long$setup <- paste0(machines_as_fixed_df_long$model, "_", machines_as_fixed_df_long$weight)
```
```{r}
joyplot_by_value <- function(df){
  df %>%
    ggplot(aes(x= estimate, y = setup, fill =setup, alpha = 0.7)) +
    facet_wrap(~value, ncol = 2) +
    geom_density_ridges() + 
    theme(legend.position = "none") + 
    xlab("score") + 
    ylab("model")
}
```

```{r}
joyplot_by_value(df = machines_as_fixed_df_long %>%
  filter(!model == "glove840B300d" | !model == "googlenews") %>%
  filter(dimension_reduction == "mean"))

joyplot_by_value(df = machines_as_fixed_df_long %>%
  filter(!model == "glove840B300d" | !model == "googlenews") %>%
  filter(dimension_reduction == "fa"))

joyplot_by_value(machines_as_fixed_df_long %>%
  filter(model == "glove840B300d" | model == "googlenews"))
```

```{r}

```


```{r}
hedonism_means_df <- machines_as_fixed_df %>% 
  select(item_ID, contains("mean")) %>%
  select(item_ID, starts_with("HEDONISM"))

machines_as_fixed_df <- machines_as_fixed_df %>% select(-contains("mean"))
```


```{r}
build_factor_loadings_df <- function(scores_df=machines_as_fixed_df, nfactors=1){
  
  # pivot to wide: each machine / value combination is a column
  # x <- values[[3]]
  
  # run factor analyses and store as list
  machine_fas <- lapply(values, function(x){
  machines_as_fixed_df %>% select(starts_with(x)) %>%
    fa(nfactors=nfactors, fm="ml")
  })
  
  # extract factor loadings as a list, format as data.table
  machine_loadings_list <- lapply(machine_fas, function(x){
    df <- x$loadings[1:8,] %>% as.data.frame()
    df$participant_ID <- rownames(df)
    rownames(df) <- NULL
    #colnames(df) <- c("loadings", "participant_ID")
    setDT(df)
    return(df)
  })

  # bind list into a single data.table
  machine_loadings_df <- rbindlist(machine_loadings_list)

  # create a new column for each value
  lapply(values, function(x){
    machine_loadings_df[grepl(x, machine_loadings_df$participant_ID), value := x]
  })

  # remove value name from participant_ID column
  for(value in values){
    machine_loadings_df$participant_ID <- gsub(paste0(value, "_"), "", machine_loadings_df$participant_ID)
  }

  machine_loadings_df$participant_ID <- gsub("_weight_lyrics", "", machine_loadings_df$participant_ID)
  machine_loadings_df$participant_ID <- gsub("mxm_", "", machine_loadings_df$participant_ID)

  return(machine_loadings_df)

}
```


```{r}
factor_1_df <- build_factor_loadings_df(nfactors=1)

factor_1_df %>%
  ggplot(aes(participant_ID, abs(.), fill = .)) +
  facet_wrap(~value, nrow=1) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_gradient2(name = "Loading", 
                       high = "blue", mid = "white", low = "red", 
                       midpoint = 0, guide = "none") +
  ylab("loading") +
  theme_bw(base_size =10) +
  theme(axis.text.x= element_text(angle = 90))
```

```{r}
show_correlation_plot <- function(df=machines_as_fixed_df, value="POWER"){
  df <- df %>% select(any_of(starts_with(value)))
  cor_mat <- cor(df)
  rownames(cor_mat) <- NULL
  colnames(cor_mat) <- NULL
  corrplot(cor_mat, title = value)
}
```

```{r}
for(value in values){
  show_correlation_plot(value=value)
}
```
```{r}
factor_2_df <- build_factor_loadings_df(nfactors=2)

factor_2_df <- factor_2_df %>% pivot_longer(cols = c("ML1", "ML2"), names_to = "factor", values_to = "eigenvalues")

factor_2_df %>%
  ggplot(aes(participant_ID, abs(eigenvalues), fill = eigenvalues)) +
  facet_wrap(value~factor) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_gradient2(name = "Loading", 
                       high = "blue", mid = "white", low = "red", 
                       midpoint = 0, guide = "none") +
  ylab("loading") +
  theme_bw(base_size =10) +
  theme(axis.text.x= element_text(angle = 90))
```

Thus it appears we have two distinct estimations: pretrained models vs. mxm data trained models. 
The weighting scheme seems to ahve had a rather small difference. 
Hedonism is the only case where the means of the mxm data trained models differ substantially from the factor scores, so they were retained. 

```{r}
machines_as_fixed_f_df <- merge(
  hedonism_means_df, 
  machines_as_fixed_df %>% select(-contains("mean")), 
  by="item_ID")

save(machines_as_fixed_f_df, file = here("VI_data_collection", "_data","intermediary_data", "machines_as_fixed_f_df.RDS"))
```

