---
title: "Timing Pilot"
output: html_document
---

# AIM:

The goal of this initial pilot was to get an initial estimate of the time it might take to complete our first draft of the `lyric preference intensity questionnaire`, and the time taken to annotate a single set of lyrics. 

We convenience sampled participants via a post on reddit.com/r/samplesize. Survey was completed on the qualtrics survey platform. 

```{r setup, include=FALSE}
library('data.table') # data manipulation
library('here')       # file logistics
library('corrplot')   # visualization
library('tidyverse')  # data manipulation
```

The actual response data is not available because it contains identifying information. 
Load de-identified versions `complete_responses_dt.RDS` and `partial_responses_dt.RDS` in the `intermediary_data` folder. 

```{r}
# path with the actual data
data_file_path <- here("I_time_pilot", "response_data")

# create a list with all of the files
# first file is the results in .sav format
# second file is the results in .csv format
# third file is the partial responses
data_files <- list.files(data_file_path)

# read in main  data file
complete_responses_dt <- fread(here(data_file_path, data_files[2])) |>
  select(-IPAddress, -LocationLatitude, -LocationLongitude)

# read in partial data file; lyric preferences are: V14:V21
partial_responses_dt <- fread(here(data_file_path, data_files[3])) |>
  select(-V6)

# remove junk on first two rows
complete_responses_dt <- complete_responses_dt[3:.N]
partial_responses_dt  <- partial_responses_dt[3:.N]

# to get question wording:
questions <- colnames(fread(here(data_file_path, data_files[2]), skip=1))

# save versions of data without identifying information
complete_responses_dt |>
  saveRDS(here("I_time_pilot", "intermediary_data", "complete_responses_dt.RDS"))

complete_responses_dt |>
  saveRDS(here("I_time_pilot", "intermediary_data", "partial_responses_dt.RDS"))

rm(data_file_path, data_files)
```

```{r}
#names of lyric preferences columns for subsetting
lyric_column_names <- c(
  "Lyric_preferences_1","Lyric_preferences_2","Lyric_preferences_3",
  "Lyric_preferences_4","Lyric_preferences_5","Lyric_preferences_6",
  "Lyric_preferences_7", "Lyric_percentage_1")

#subset data table
lyrics_preferences_dt <- complete_responses_dt[, lyric_column_names, with=FALSE]

rm(lyric_column_names)
```

```{r}
# names of lyric preferences columns for subsetting
partial_lyric_column_names <- c("V14", "V15", "V16", "V17", "V18", "V19", "V20", "V21")

# subset data table
partial_lyrics_preferences_dt <- partial_responses_dt[, partial_lyric_column_names, with=FALSE]

# rename columns
colnames(partial_lyrics_preferences_dt) <- c("L1", "L2", "L3", "L4", "L5", "L6", "L7", "L8")

# convert all columns to numeric
partial_lyrics_preferences_dt <- sapply(partial_lyrics_preferences_dt, as.numeric)

rm(partial_lyric_column_names)
```

A first look at the lyric preferences questionnaire:

```{r}
#function to recode values to numbers
to_numbers <- function(x) {
  if(x ==  "Strongly disagree"){x = as.numeric(1)
  } else if (x== "Somewhat disagree"){x = as.numeric(2)
  } else if (x== "Neither agree nor disagree"){x = as.numeric(3)
  } else if (x=="Somewhat agree"){x = as.numeric(4)
  } else if (x=="Strongly agree"){x = as.numeric(5)
}}

#execute recode function on relevant rows
lyrics_preferences_dt <- lyrics_preferences_dt[, .(
  L1 = lapply(Lyric_preferences_1, to_numbers), 
  L2 = lapply(Lyric_preferences_2, to_numbers), 
  L3 = lapply(Lyric_preferences_3, to_numbers),
  L4 = lapply(Lyric_preferences_4, to_numbers), 
  L5 = lapply(Lyric_preferences_5, to_numbers), 
  L6 = lapply(Lyric_preferences_6, to_numbers), 
  L7 = lapply(Lyric_preferences_7, to_numbers),
  L8 = Lyric_percentage_1)][, lapply(.SD, as.numeric)]

colnames(lyrics_preferences_dt) <- c("L1", "L2", "L3", "L4", "L5", "L6", "L7", "L8")

# merge two tables
lyrics_preferences_dt <- rbind(lyrics_preferences_dt, partial_lyrics_preferences_dt)

rm(partial_lyrics_preferences_dt)
```


```{r}
#extract actual wording of questions
lyrics_questions <- questions[19:26]
lyrics_questions <- gsub("Please indicate how much you agree with the following statements about your music preferences: - ", "", lyrics_questions)
lyrics_questions

#plot correlation matrix
lyrics_preferences_dt |>
  as.matrix() |>
  cor(use = "complete.obs") |>
  corrplot(
      method = "color", 
      type = "full", 
      addCoef.col = "black", 
      number.cex = 1)

#save data for further analysis
path <- here("I_time_pilot", "intermediary_data", "lyrics_preferences.RDS")
saveRDS(lyrics_preferences_dt, path)

rm(lyrics_preferences_dt, partial_lyrics_preferences_dt, partial_responses_dt, lyrics_questions, to_numbers, questions, path)
```

I used only complete responses to estimate the amount of time it took to complete the items in the survey. 

```{r}
clicks_dt <- complete_responses_dt

# remove spaces from column names
setnames(clicks_dt, make.names(colnames(clicks_dt), unique=TRUE))

# logical vector with all the 'click' columns and response ID
clicks <- grepl("First.Click|Last.Click|ResponseId", colnames(clicks_dt)) 

# filter data.table by clicks vector
# make all values numeric
# compute the mean time taken for each component
summary_dt <- rbind(
clicks_dt[, ..clicks][, lapply(.SD, as.numeric)][, .(
  lyric_prefs = mean(Lyric_questions_time_Last.Click - Lyric_questions_time_First.Click, na.rm=TRUE) / 60, 
  TRKMVRW128F4292AF7 = mean(TRKMVRW128F4292AF7.m_Last.Click - TRKMVRW128F4292AF7.m_First.Click) / 60, 
  TRAKHVT128F4299C0C = mean(TRAKHVT128F4299C0C.m_Last.Click - TRAKHVT128F4299C0C.m_First.Click) / 60, 
  TRIOIJF12903CF5B9D = mean(TRIOIJF12903CF5B9D.m_Last.Click - TRIOIJF12903CF5B9D.m_First.Click) / 60, 
  TRFUXHO128F930BBFE = mean(TRFUXHO128F930BBFE.m_Last.Click - TRFUXHO128F930BBFE.m_First.Click) / 60
)],

# compute the sd of time taken for each component
clicks_dt[, ..clicks][, lapply(.SD, as.numeric)][, .(
  lyric_prefs = sd(Lyric_questions_time_Last.Click - Lyric_questions_time_First.Click, na.rm=TRUE) / 60, 
  TRKMVRW128F4292AF7 = sd(TRKMVRW128F4292AF7.m_Last.Click - TRKMVRW128F4292AF7.m_First.Click) / 60, 
  TRAKHVT128F4299C0C = sd(TRAKHVT128F4299C0C.m_Last.Click - TRAKHVT128F4299C0C.m_First.Click) / 60, 
  TRIOIJF12903CF5B9D = sd(TRIOIJF12903CF5B9D.m_Last.Click - TRIOIJF12903CF5B9D.m_First.Click) / 60, 
  TRFUXHO128F930BBFE = sd(TRFUXHO128F930BBFE.m_Last.Click - TRFUXHO128F930BBFE.m_First.Click) / 60
)])

summary_dt[, stat := c("mean", "sd")] #create summary data table
```

```{r}
#identify columns that contain the string that identifies items
lyric_items <- grepl("TRKMVRW128F4292AF7.1|TRAKHVT128F4299C0C.1|TRIOIJF12903CF5B9D.1|TRFUXHO128F930BBFE.1", colnames(complete_responses_dt))

#recreate all the column names
TRKMVRW128F4292AF7 <- paste0("TRKMVRW128F4292AF7.1_", seq(1:10))
TRAKHVT128F4299C0C <- paste0("TRAKHVT128F4299C0C.1_", seq(1:10))
TRIOIJF12903CF5B9D <- paste0("TRIOIJF12903CF5B9D.1_", seq(1:10))
TRFUXHO128F930BBFE <- paste0("TRFUXHO128F930BBFE.1_", seq(1:10))

#take the responses data table, and filter by lyric items
#then make all the values in the cells numeric
na_dt <- complete_responses_dt[, ..lyric_items][, lapply(.SD, as.numeric)]

#single row data table with the number of NAs
na_dt <- na_dt[, .(
  lyric_prefs = NA,
  TRKMVRW128F4292AF7 = sum(is.na(na_dt[, ..TRKMVRW128F4292AF7])) / 120, 
  TRAKHVT128F4299C0C = sum(is.na(na_dt[, ..TRAKHVT128F4299C0C])) / 120, 
  TRIOIJF12903CF5B9D = sum(is.na(na_dt[, ..TRIOIJF12903CF5B9D])) / 120, 
  TRFUXHO128F930BBFE = sum(is.na(na_dt[, ..TRFUXHO128F930BBFE])) / 120,
  stat = "prop_missing"
)]

#bind NA data table to the summary data table
summary_dt <- rbind(summary_dt, na_dt) 
```

```{r}
#take the responses data table, and filter by lyric items
#then make all the values in the cells numeric
zero_dt <- complete_responses_dt[, ..lyric_items][, lapply(.SD, as.numeric)]

#single row data table with the number of 0s
zero_dt <- zero_dt[, .(
  lyric_prefs = NA,
  TRKMVRW128F4292AF7 = sum(colSums(as.matrix(zero_dt[,..TRKMVRW128F4292AF7])==0, na.rm=TRUE)) / 120, 
  TRAKHVT128F4299C0C = sum(colSums(as.matrix(zero_dt[,..TRAKHVT128F4299C0C])==0, na.rm=TRUE)) / 120, 
  TRIOIJF12903CF5B9D = sum(colSums(as.matrix(zero_dt[,..TRIOIJF12903CF5B9D])==0, na.rm=TRUE)) / 120, 
  TRFUXHO128F930BBFE = sum(colSums(as.matrix(zero_dt[,..TRFUXHO128F930BBFE])==0, na.rm=TRUE)) / 120,
  stat = "prop_zero"

)]

#bind to summary data table
summary_dt <- rbind(summary_dt, zero_dt) 

colnames(summary_dt) <- c("lyric preferences", "song 1", "song 2", "song 3", "song 4", "statistic")
summary_dt <- summary_dt |> select(statistic, everything())
```

Participants took between 1 and 2 minutes to complete the questions for each set of lyrics, with the amount of time decreasing as they progressed through the lyrics. 

```{r}
path <- here("I_time_pilot", "intermediary_data", "time_pilot_summary.RDS")

summary_dt |> saveRDS(file = path)

rm(clicks_dt, complete_responses_dt, na_dt, zero_dt, clicks, lyric_items, TRAKHVT128F4299C0C, TRFUXHO128F930BBFE, TRIOIJF12903CF5B9D, TRKMVRW128F4292AF7)
```

