---
title: "estimate correlations, single machines model, with Mplus"
---

This notebook estimates the correlations between the participant ratings, and the individual machine setups using Mplus. 

```{r}
library('data.table')      # data manipulation
library('here')            # file logistics
library('tidyverse')       # data manipulation
library("MplusAutomation") # mplus integration
library("ggplot2")         # visualization
library("hash")            # dictionaries / hash tables
library("stringr")         # string manipulation


values <- c("POWER", "ACHIEVEMENT", "HEDONISM", 
            "STIMULATION", "SELF", "UNIVERSALISM", 
            "BENEVOLENCE", "TRADITION", 
            "CONFORMITY", "SECURITY")

# load dataset
source(here("VI_data_collection", "7_participant_machine_correlation", "0_data_set_merge_1.0.R"))
```


```{r}
# remove columns not in analysis
cols <- c("writer", "familiar", "confidence")

df <- responses_machines_as_fixed_df %>% 
  # columns not in analysis
  select(-all_of(cols)) %>% 
  # remove mean columns
  select(-contains("mean")) 

# make dataframe of columns names
col_names <- colnames(df) %>% as.data.frame()

#col_names$mplus_name <- c("row_number", "item_ID", "participant_ID",
col_names$mplus_name <- c("item_ID", "row_number", "participant_ID",
    paste0("y", seq(1,10)), 
    paste0("z", seq(1, 80))
  ) %>% toupper()
colnames(col_names) <- c("og_names", "mplus_name")

names <- hash(keys = col_names$mplus_name, values = col_names$og_names)

rm(responses_machines_as_fixed_df, cols)
```

```{r}
# set column names for mplus
colnames(df) <- c("row_number", "item_ID", "participant_ID",
  paste0("y", seq(1,10)), 
  paste0("z", seq(1, 80)))

df <- as.data.frame(df)
```

```{r}
# load mplus model
source(here("VI_data_collection", "7_participant_machine_correlation", "_mplus_models", "1_correlation_single_machines_model_1.1.R"))
```


```{r}
 fit <- mplusModeler(mplus_model, 
                     modelout = "VI_data_collection/_data/mplus_output/sin.inp", 
                     hashfilename = TRUE,
                     check = TRUE,
                     run=TRUE)

# print errors and warnings
 fit$results$errors
 fit$results$warnings
 
rm(fit, mplus_model)
```
The first type of standardization is shown under the heading StdYX in the output.  StdYX uses the variances of the continuous latent variables as well as the variances of the background and outcome variables for standardization.  The StdYX standardization is the one used in the linear regression of y on x,

bStdYX = b*SD(x)/SD(y),

where b is the unstandardized linear regression coefficient, SD(x) is the sample standard deviation of x, and SD(y) is the model estimated standard deviation of y.  The standardized coefficient bStdYX is interpreted as the change in y in y standard deviation units for a standard deviation change in x.

```{r}
out <-readModels(target = here("VI_data_collection/_data/mplus_output/sin.out"))

output_df <- out$parameters$ci.stdyx.standardized

rm(out)
```

```{r}
corr_df <- output_df %>% 
  filter(grepl('.WITH', paramHeader)) %>%
  select(paramHeader, param, low2.5, est, up2.5) 

corr_df$model <- col_names$og_names[match(corr_df$param,col_names$mplus_name)]
```

```{r}
lapply(values, function(x){
  corr_df %>%
    filter(grepl(x, model)) %>%
      ggplot(aes(x = est, y = model, color = param)) +
        geom_point() +
        geom_errorbar(aes(xmin=`low2.5`, xmax=`up2.5`)) +
        scale_x_continuous(limits = c(-.5, +.5)) + 
        theme_minimal() +
        theme(legend.position = "none")
})
```


```{r}
model_names <- list("far_idf_", "far_uniform_",
                "split_idf_", "split_uniform_",	
                "glove840B300d_idf_", "glove840B300d_uniform_", 
                "googlenews_idf_", "googlenews_uniform_")

lapply(model_names, function(x){
    corr_df %>%
    filter(grepl(x, model)) %>%
      ggplot(aes(x = est, y = model, color = param)) +
        geom_point() +
        geom_errorbar(aes(xmin=`low2.5`, xmax=`up2.5`)) +
        scale_x_continuous(limits = c(-.5, +.5)) + 
        theme_minimal() +
        theme(legend.position = "none")
})
```

```{r}
corr_df$value <- word(corr_df$model, 1, sep = "\\_")
corr_df$setup <- word(corr_df$model, 2, -1, sep = "\\_")
corr_df$setup <- gsub("_lyrics", "", corr_df$setup)
corr_df$setup <- gsub("_weight", "", corr_df$setup)
```

```{r}
    corr_df %>%
      ggplot(aes(x = est, y = setup, color = param)) +
        facet_wrap(~value, ncol=2) +
        geom_point() +
        geom_errorbar(aes(xmin=`low2.5`, xmax=`up2.5`)) +
        scale_x_continuous(limits = c(-.5, +.5)) + 
        theme_minimal() +
        theme(legend.position = "none")
```

```{r}
    corr_df %>%
      ggplot(aes(x = est, y = value, color = param)) +
        facet_wrap(~setup, ncol=2) +
        geom_point() +
        geom_errorbar(aes(xmin=`low2.5`, xmax=`up2.5`)) +
        scale_x_continuous(limits = c(-.5, +.5)) + 
        theme_minimal() +
        theme(legend.position = "none")
```

