---
title: "time_per_item"
author: "andrew demetriou"
date: "26/04/2023"
---

#AIM

The aim of this notebook is to examine how long it took to answer the questions for each song (item_ID), and the questions on the lyrics preference intensity questionnaire.
The dataframe used here is not in the public repo. It includes both complete and partial responses for which individuals indicated consent on the consent form. 


```{r setup, include=FALSE}
library('here')       # file logistics
library('data.table') # data manipulation
library('dplyr')      # data manipulation
library('ggplot2')    # data visualization
library('gridExtra')  # data visualization
library('naniar')     # visualizing missingness
library('lubridate')  # minute / second functions
library('forcats')    # ordering factors
library('lme4')       # linear mixed effects model 
library('lmerTest')   # p values for lme4

options(scipen=999)
```

```{r}
# load dataset
load(here("VI_data_collection_november_2022", "intermediary_data", "working_df.RDS"))

# create a working dataset
responses_dt <- as.data.table(working_df)

rm(working_df)
```

Less than 20% of respondents left feedback. 

```{r}
# examine missingess of variables
missingness_dt <- responses_dt[,participant_ID:Finished] |> unique()

# make a summary table
missingness_dt |> 
  group_by(Finished) |>
  select(-starts_with("Lyric")) |>
  miss_var_summary()

# demographic data for surveys that were not completed is mostly unavailable

rm(missingness_dt)
```


Median age was 40, range 18-77, with 67 missing values. 

```{r}
# coerce to numeric
responses_dt$Age <- as.numeric(responses_dt$Age)

age_dt <- responses_dt |> 
  select(participant_ID, Age) |>
  base::unique()

# median age is 40
median(age_dt$Age, na.rm=TRUE)

# plot histogram
age_dt |>
  ggplot(aes(x=Age)) +
  geom_histogram(bins=50) +
  geom_vline(xintercept = 40, color="red") +
  geom_vline(xintercept = 18, color="blue") +
  geom_vline(xintercept = 77, color="blue") +
  annotate(geom="text", x=46, y=30, label="Median = 40") +
  annotate(geom="text", x=22, y=30, label="Min = 18") +
  annotate(geom="text", x=72, y=30, label="Max = 77") +
  xlab("Respondent Age") +
  ylab("") +
  theme_minimal()

rm(age_dt)
```

Both prolific and qualtrics have their own measures of time taken for the entire survey. Here I take a look at each of them. 

```{r}
#rename qualtrics duration column
responses_dt$duration <- as.numeric(responses_dt$`Duration (in seconds)`) 
responses_dt$`Time taken` <- as.numeric(responses_dt$`Time taken`)

#define columns to plot
columns <- c("duration", "participant_ID", "Time taken", "Sex", "Age", "Finished")

# create missingness matrix
duration_dt <- responses_dt |> 
  select(all_of(columns)) |> 
  base::unique() |>
  bind_shadow()

rm(columns)
```

Qualtrics recorded some very long durations; the longest is 100521 seconds. These appear to be unfinished surveys that are not reflected in the prolific data. 

```{r}
#plot survey duration according to qualtrics
duration_plot <- duration_dt %>%
  ggplot(., aes(x=duration, y=Age, color=Finished)) +
  geom_miss_point() +
  xlab("Time (in seconds): Qualtrics") +
  theme_minimal()

#plot survey duration according to prolific
time_taken_plot <- duration_dt %>% 
  ggplot(., aes(x=`Time taken`, y=Age, color=Finished)) +
  geom_miss_point() +
  xlab("Time (in seconds): Prolific") +
  theme_minimal()

grid.arrange(duration_plot, time_taken_plot, ncol=1)

rm(duration_dt, duration_plot, time_taken_plot)
```

Those that didn't finish the survey don't appear in the prolific data. 
There doesn't seem to be a pattern that relates age to the time it took to complete the survey. 

```{r}
# create df with single row for each participant
time_taken_df <- responses_dt %>% 
  select(participant_ID, `Time taken`) %>%
  base::unique() 

# create a readout with minutes and seconds
time_taken_df <- time_taken_df |>
  mutate(sec = `Time taken` %% 60, 
         min = round(`Time taken` / 60), 
         minsec = paste(min, sec), 
         minsec = ms(minsec)) |>
  select(-sec, -min)

median(time_taken_df$`Time taken`/60, na.rm=T)

# plot time taken
time_taken_df %>% ggplot(aes(x=`Time taken`/60)) +
  geom_histogram() +
  geom_vline(xintercept = 33.86667, color="red") +
  annotate(geom="text", x=42, y=50, label="Median = 33.86") +
  xlab("Time (in minutes): Prolific") +
  ylab("") +
  theme_minimal()

rm(time_taken_df)
```



```{r}
# create an outliers dataset to examine participant responses
outliers_dt <- responses_dt[`Time taken`>5000 | duration>5000]

# responses look legit

rm(outliers_dt)
```

Here I select rows that have the timing information for each of the questions. Specifically, I select rows with that contain a) the first click on the page, and b) the last click on the page. I then subtract the first click time from the last click time to get the time spent on the page. 

According to: https://www.qualtrics.com/support/survey-platform/survey-module/editing-questions/question-types-guide/advanced/timing/, the time is in seconds. 


```{r}
# logical vector of matches to: 
# 1) either first or last click for a question
# 2) participant ID
# 3) the song id
clicks <- grepl("m_First Click|m_Last Click|participant_ID|item_ID|Age|Sex", colnames(responses_dt)) 

# subset data based on clicks vector
clicks_dt <- responses_dt[, ..clicks]

# time taken per song
clicks_dt <- clicks_dt[, diff := (as.numeric(`m_Last Click`) - as.numeric(`m_First Click`))]

rm(clicks)
```

Visualize differences distribution of time taken per song. 
Here I take the 40 songs with the largest standard deviation and plot them against the songs with the lowest standard deviation. 

```{r}
# create a df of median time taken for each song
time_taken_table <- clicks_dt |> 
  group_by(item_ID) |>
  summarize(mean = mean(diff, na.rm=T), median = median(diff, na.rm=T), sd = sd(diff, na.rm=T)) |>
  # descending order by standard deviation
  arrange(desc(sd))

# the 40 songs with the largest standard deviation
item_IDs <- head(time_taken_table$item_ID, 20)

# plot largest sd songs
long_diff_plot <- clicks_dt[item_ID %in% item_IDs] %>%
  ggplot(., aes(x=diff, y=item_ID, color = item_ID)) +
  geom_boxplot(show.legend=FALSE) +
  theme_minimal() +
  xlab("Time taken (In seconds)") + 
  xlim(0, 750)

# plot smallest sd songs
item_IDs <- tail(time_taken_table$item_ID, 20)
short_diff_plot <- clicks_dt[item_ID %in% item_IDs] %>%
  ggplot(., aes(x=diff, y=item_ID, color = item_ID)) +
  geom_boxplot(show.legend=FALSE) +
  theme_minimal() +
  xlab("Time taken (In seconds)") + 
  xlim(0, 750)

grid.arrange(long_diff_plot, short_diff_plot, ncol=1)

rm(time_taken_table, long_diff_plot, short_diff_plot)
```

Differences are quite substantial. 

```{r}
setDT(not_simple_medium_alpha_2021_full_2_2)
long_sd_df <- not_simple_medium_alpha_2021_full_2_2[mxm_id %in% item_IDs]
```


```{r}
# an exploratory random effects model looking at age and time per question block
model <- lmer(diff ~ (1|item_ID) + Age, data = clicks_dt)

summary(model)

confint(model, level = 0.95)

# even if significant, the effect appears to be very small
# .21, (95%CI: .16, .26) seconds
# age range 18 - 77

rm(model)
```


Although there's no reason to expect sex differences I plot them anyway.
There doesn't seem to be any sex differences. 

```{r}
clicks_dt[diff<900] %>% 
  filter(!is.na(Sex)) %>%
  ggplot(aes(x=diff, color=Sex, fill=Sex)) +
  geom_histogram() +
  facet_wrap(~Sex) +
  xlab("Time (in seconds)") +
  ylab("") +
  theme_minimal()

clicks_dt[diff<900] %>% ggplot(aes(x=diff)) +
  geom_histogram(bins = 30) +
  geom_vline(aes(xintercept = median(diff)), color="red") +
  annotate(geom="text", x=200, y=4000, label="Median = 31.8") +
  xlab("Time (in seconds)") +
  ylab("") +
  theme_minimal()
```

```{r}
clicks_dt |>
  ggplot(aes(x=diff, y=Age, color=Sex)) +
  geom_point(alpha = 0.1) +
  theme_minimal() +
  xlab("Time (in seconds)")
```

Here I filter out the initial lyrics questionnaire and estimate central tendency of the time taken per lyric. 

```{r}
no_outliers_removed <- clicks_dt |>
  summarize(median = median(diff, na.rm=T), range_lower = range(diff, na.rm=T)[1], range_upper = range(diff, na.rm=T)[2], mean = mean(diff, na.rm=T), sd = sd(diff, na.rm=T))

outliers_less_than_200 <- clicks_dt[diff<200] |>
  summarize(median = median(diff, na.rm=T), range_lower = range(diff, na.rm=T)[1], range_upper = range(diff, na.rm=T)[2], mean = mean(diff, na.rm=T), sd = sd(diff, na.rm=T))

central_tendency <- rbind(no_outliers_removed, outliers_less_than_200)
central_tendency$outliers <- c("no outliers removed", "outliers set at < 200")
central_tendency <- central_tendency |> select(outliers, everything())
central_tendency

rm(no_outliers_removed, outliers_less_than_200, central_tendency, clicks_dt)
```

Thus, I don't expect differences due to age or sex in the time taken to complete a question block. 
Participants took a median 25.7 seconds to complete each question block, but there is an extremely large range. 

Here I examine how long it took to complete the Lyric preference intensity questionnaire:

```{r}
# subset questions
lyric_questions_clicks_dt <- responses_dt |>
  select(participant_ID, Finished, starts_with("Lyric_questions_time_")) |>
  unique()

# convert to numeric
lyric_questions_clicks_dt$`Lyric_questions_time_Last Click` <- as.numeric(lyric_questions_clicks_dt$`Lyric_questions_time_Last Click`)
lyric_questions_clicks_dt$`Lyric_questions_time_First Click` <- as.numeric(lyric_questions_clicks_dt$`Lyric_questions_time_First Click`)

# compute difference
lyric_questions_clicks_dt$diff = lyric_questions_clicks_dt$`Lyric_questions_time_Last Click` - lyric_questions_clicks_dt$`Lyric_questions_time_First Click`
```

```{r}
finished_isna <- is.na(lyric_questions_clicks_dt$Finished )
lyric_questions_clicks_dt <- lyric_questions_clicks_dt[! finished_isna]
```

A number of rows look like the data was poorly recorded; time taken evaluates to 0, and only 2-3 clicks on the page were recorded. 

There are also 3 extreme values, two of which show over 1500 seconds, and one showing over 900 seconds, among participants who completed the survey. 

```{r}
lyric_questions_clicks_dt |>
  ggplot(aes(x=diff, y=Finished, color = Finished)) +
  geom_point(alpha = .5, position = 'jitter') + 
  xlab("Time (in seconds)") +
  ylab("") +
  theme_minimal()
```
```{r}
lyric_questions_clicks_dt[diff<750] |>
  summarize(median = median(diff), range_lower = range(diff)[1], range_upper=range(diff)[2])

lyric_questions_clicks_dt[diff<750] |>
  ggplot(aes(diff)) +
  geom_histogram(bins=40) +
  geom_vline(xintercept=82.461, color = "red") +
  annotate(geom="text", x=190, y=100, label="Median = 82.461") +
  xlab("Time (in seconds)") +
  ylab("") +
  theme_minimal()
```



