---
title: "factor analysis lyric preference intensity questionnaire"
---

```{r}
library('data.table') # data manipulation
library('dplyr')      # data manipulation
library('tidyverse')  # data manipulation
library('here')       # file logistics
library('ggplot2')    # visualization
library('corrplot')   # visualize correlations
library('Hmisc')      # visualization
library('psych')      # factor analysis
library('mirt')       # item response theory
library('shiny')      # interactive visualization

options(scipen=999)
theme_set(theme_minimal())
```

```{r}
# load dataset
load(here("VI_data_collection_november_2022", "intermediary_data", "public_df.RDS"))

# create a working dataset
responses_dt <- as.data.table(public_df)

rm(public_df)
```


```{r}
# subset lyrics df
lyric_preferences_dt <- responses_dt %>%
  select(
    participant_ID, 
    subjective,
    all_of(starts_with("Lyric"))) %>%
  unique()

# rename columns
colnames(lyric_preferences_dt) <- gsub("yric_preferences", "", colnames(lyric_preferences_dt))
names(lyric_preferences_dt)[names(lyric_preferences_dt) == 'Lyric_percentage_1'] <- 'L_16'

rm(responses_dt)
```

```{r}
# subset lyric preferences items
lyrics_questionnaire <- lyric_preferences_dt[,3:18]

#ensure that only complete cases are included
lyrics_questionnaire <- lyrics_questionnaire[complete.cases(lyrics_questionnaire),]

# convert all columns to numeric
lyrics_questionnaire <- mutate_all(lyrics_questionnaire, function(x) as.numeric(as.character(x)))
```

```{r}
# function to plot histogram for each question
histogram_plot_function <- function(column){
  
  ggplot(lyrics_questionnaire, aes(x=column)) +
    geom_histogram(bins=9) +
    scale_y_continuous(limits = c(0,300))
}

list_of_histograms <- lapply(lyrics_questionnaire, histogram_plot_function)
```

As per https://bookdown.org/bean_jerry/using_r_for_social_work_research/item-response-theory.html:

The better fitting model shows and SRMSR of .085, and an RMSEA of .08.

```{r}
# reset scale
lyrics_questionnaire$L_16 <- lyrics_questionnaire$L_16 / 20

#estimate Graded Response model (GRM)
mod_1 <- (mirt(lyrics_questionnaire, 1, verbose = FALSE, itemtype = 'graded', SE = TRUE))

#model fit for response models with ordinal data
M2(mod_1, type = "C2", calcNULL = FALSE)


#estimate Generalized Partial Credit Model
mod_2 <- (mirt(lyrics_questionnaire, 1, verbose = FALSE, itemtype = 'gpcm', SE = TRUE))

#model fit
M2(mod_2, type = "C2", calcNULL = FALSE)

# model 2 has a better fit: RMSEA = .08, SRMSR =.085
mod <- mod_2
rm(mod_1, mod_2)
```

"The mirt implementation of S-X2 computes an RMSEA value which can be used to assess degree of item fit. Values less than .06 are considered evidence of adequate fit."

It appears the model fits all the items well, as all RMSEA - S-X2 values are below .06. 

```{r}
itemfit(mod)
```


Items 1, 2, 5, and 16 have extremely flat lines. 

```{r}
#plot(mod, type='trace', which.item = c(1,2,3,4,5,6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16), facet_items=T, 
#     as.table = TRUE, auto.key=list(points=F, lines=T, columns=4, space = 'top', cex = .8), 
#              theta_lim = c(-3, 3), 
#     main = "")

plot(mod, type='infotrace', which.item = c(1,2,3,4,5,6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16), facet_items=T, 
     as.table = TRUE, auto.key=list(points=F, lines=T, columns=4, space = 'top', cex = .8), 
              theta_lim = c(-3, 3), 
     main = "")
```


```{r}
# scree plot: 4 factors
fa.parallel(lyrics_questionnaire)

# adjusted scale: 3 factors
lyrics_questionnaire %>%
  select(-L_1, -L_2, -L_5, -L_16) %>%
  fa.parallel()
```

```{r}
# alpha is .79, (95% CI: .77, .82)
lyrics_alpha <- alpha(lyrics_questionnaire, n.iter = 500, check.keys = T)
lyrics_alpha$boot.ci

# alpha improces to .83, (95% CI: .81, .85)
lyrics_alpha <- lyrics_questionnaire %>% 
    select(-L_1, -L_2, -L_5, -L_16) %>%
    alpha(., n.iter=500, check.keys=T)
lyrics_alpha$boot.ci

rm(lyrics_alpha)
```

```{r}
# fa diagrams confirm that questions 1, 2, 5, and 16 may not be very informative

  lapply(seq(1,4), function(x){
    fa(lyrics_questionnaire, rotate = "varimax", fm="ml", nfactors = x) %>%
  fa.diagram()
})
```

```{r}
# model gets closer to acceptable parameters: RMSEA = .075, SRMSR = .066

#estimate Generalized Partial Credit Model
mod <- (mirt(lyrics_questionnaire %>% select(-L_1, -L_2, -L_5, -L_16), 1, verbose = FALSE, itemtype = 'gpcm', SE = TRUE))

#model fit
M2(mod, type = "C2", calcNULL = FALSE)
```
Scale information / test infomation and conditional standard errors:

Our final scale did not do a good job of measuring values at the extremes:

```{r}
plot(mod, type = 'infoSE', theta_lim = c(-3, 3), 
     main="")
```

```{r}
factor_analysis <- fa(lyrics_questionnaire, 3, rotate = "varimax", fm="ml")
factor_scores   <- factor.scores(lyrics_questionnaire, factor_analysis)
factor_scores   <- factor_scores$scores
lyric_preferences_dt <- cbind(lyric_preferences_dt, factor_scores)

rm(factor_analysis, factor_scores)
```

```{r}
#estimate p values for each correlation in a matrix
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
```

A very small correlation between the first factor and subjectivity ratings, likely spurious. 
Also moderate correlations between the factors scores. 

```{r}
corr_df <- lyric_preferences_dt %>% select(subjective, ML1, ML2, ML3) %>% 
  mutate_all(., function(x) as.numeric(as.character(x))) %>%
  filter(!is.na(subjective)) 

#corr_df %>% cor()

p.mat <- cor.mtest(corr_df)

corrplot(cor(corr_df), p.mat = p.mat, sig.level=0.05, type = "lower", diag = F)

rm(corr_df, p.mat, cor.mtest)
```



```{r}
all_lyrics_questions <- c(
  "I prefer music that contains lyrics, as opposed to music that does not", 
  "I only pay attention to the lyrics of songs or artists that I like.", 
  "I always pay attention to the lyrics of a song, if the song has them", 
  "I enjoy learning about song lyrics and their meaning, for example by reading  blogs and forums or listening to artist interviews", 
  "If a song has lyrics that I don't like for any reason, I don't listen to it", 
  "If I am not sure about the lyrics of a song, I search them on the internet", 
  "I contribute to online resources on lyrics, for example forums, or on platforms where I can contribute lyric transcriptions", 
  "I memorize the lyrics to the songs I listen to", 
  "I write my own song lyrics", 
  "I post excerpts of song lyrics online, e.g. on social media", 
  "I discuss song lyrics with my friends", 
  "I come up with alternate versions of song lyrics that I find entertaining, i.e. song parodies", 
  "I ponder the meaning of song lyrics", 
  "I quote lyrics often in conversation", 
  "I read and/or write poetry", 
  "What percentage of your music library do you think contains songs with lyrics?" 
)
```

```{r}
# question lists
four_factor_questions <- 
  list(all_lyrics_questions[c(10, 14, 7, 11, 15, 12)],
     all_lyrics_questions[c(13, 4, 6, 3, 2)],
     all_lyrics_questions[9],
     all_lyrics_questions[c(1, 16, 8)])

three_factor_questions <- 
  list(all_lyrics_questions[c(10,7,14,11,15,9,12)], 
       all_lyrics_questions[c(13, 4, 6, 3, 2)], 
       all_lyrics_questions[c(1,16,8)])

two_factor_questions <- 
  list(all_lyrics_questions[c(11,10,14,7,15,12,9)], 
       all_lyrics_questions[c(3,16,4,13,8,6,1)])

one_factor_questions <- 
  list(all_lyrics_questions[c(11,4,14,13,10,15,6,3,8,7,12,9)])

question_lists <- list(one_factor_questions, two_factor_questions, three_factor_questions, four_factor_questions)

rm(one_factor_questions, two_factor_questions, three_factor_questions, four_factor_questions)
```



```{r}
ui <- fluidPage(
  titlePanel("Lyric Questions"),
  
  # drop down menu to display the values
  numericInput('question', 'select question number', value=1, min=1, step=1, max=16), 
  
  # print out specific question
  uiOutput('question_text'),
  
  #plot histogram of question responses
  plotOutput('histogram'),
  
  # print out all lyrics questions
  verbatimTextOutput('all_questions'),
  
  numericInput('factors', 'select number of factors', value=1, min=1, step=1, max=4),
  
  #correlation plot
  plotOutput('fa_plot')
)

server <- function(input, output, session){
  
  # retrieve input list element
  output$question_text <- renderUI(all_lyrics_questions[input$question])
  
  # show question list
  output$all_questions <- renderPrint(question_lists[[input$factors]])
  
  # plot input item histogram
  output$histogram <- renderPlot({
    list_of_histograms[input$question] 
  })
  
  # correlation plot of lyrics responses
  output$fa_plot <- renderPlot({
    #list_of_diagrams[input$factors]
        fa(lyrics_questionnaire, rotate = "varimax", fm="ml", nfactors = input$factors) %>%
  fa.diagram()
  })
  
  
}
shinyApp(ui=ui, server=server)
```





