---
title: "helper notebook for qualtrics data entry"
author: "Andrew M. Demetriou"
date: '2022-11-17'
---

# AIM: To faciliate copy+paste data entry to Qualtrics

This notebook was used to display the mxm id of each song along with the raw lyrics, to make copy and pasting the data into a Qualtrics Survey. 

Note: the original plan was to use formR. However, the surver at Gottingen could not be verified as GDPR compliant by our data stewards, and our attempt to set up a formR server at TU Delft failed. Thus we defaulted to Qualtrics. 

```{r}
library("here")
library("data.table")
library("tidyverse")
library("knitr")
```

```{r}
# file names: SET THESE TO THE CURRENT FILES
lyrics_file    <- "not_simple_medium_alpha_2021_full_2_2.csv"
selection_file <- "not_simple_medium_alpha_2021_full_10w (2).csv"

# paths with data and template
primary_data_file_path <- here("IV_survey_builder", "primary_data_")
template_file_path <- here("IV_survey_builder", "template_files_")
```

```{r}
# lyrics datasets
all_lyrics_dt     <- fread(here(primary_data_file_path, lyrics_file)) %>% select(mxm_id, lyrics_body)

# manual selection process
selected_dt <- fread(here(primary_data_file_path, selection_file)) %>% 
  select(mxm_id, Keep) %>% 
  # these are the four identifiers from the lyric rating process
  filter(Keep == "O" | Keep == "o" | Keep == "m" | Keep == "M" | Keep == "a" | Keep == "A")

# making a single df
selected_dt <- merge(selected_dt, all_lyrics_dt, be=mxm_id) |>
  select(mxm_id, lyrics_body)

rm(all_lyrics_dt, lyrics_file, primary_data_file_path, template_file_path, selection_file)
```

```{r}
row_number <- 361

kable(selected_dt[row_number,2])

kable(selected_dt[row_number,1])

#-t
#-w
#-f
#-t
#-1
#-m
#-c
#-r
```



