---
title: "manual_lyric_adjustment"
author: "andrew demetriou"
date: "4/21/2022"
---

#AIM

This notebook details manual changes to the raw lyrics we downloaded from musixmatch. They run in reverse order, with the most recent changes shown first, and earlier changes shown later. 


```{r setup}
library("here")
library("data.table")
library("dplyr")
```

Here we document manual changes to song lyrics based on discussions among the research group, resulting in primary data __version 2_2__. These manual changes include:
* the minor adjustment of non-English characters, 
* the removal of time-stamps, 
* the removal of artist and/or song names, and 
* the removal of section indications (e.g. 'verse', 'chorus' etc.). 

At this time we found 18 songs that could be manually adjusted for inclusion in our study, denoted by the 'M' indicator in our spreadsheet. 

```{r import_files_2.2}
# path with data
data_file_path <- here("IV_survey_builder", "primary_data_")
lyrics_file    <- "not_simple_medium_alpha_2021_full_2.csv"
selection_file <- "not_simple_medium_alpha_2021_full_10w (2).csv"

# manual selection process
selected_dt <- fread(here(data_file_path, selection_file)) %>% 
  select(mxm_id, Keep) %>% 
  # these are the four identifiers from the lyric rating process
  filter(Keep == "M" | Keep == "m")

# make smaller dt with only adjustable songs
songs <- selected_dt[,1]

# lyrics datasets
all_lyrics_dt     <- fread(here(data_file_path, lyrics_file)) %>% select(mxm_id, lyrics_body)
```

We remove artist names from the first song:

```{r first_song_2.2}
# select first song
song <- as.character(songs[1,1])

#view complete lyrics
knitr::kable(all_lyrics_dt[mxm_id %in% song][,2])

# replace characters
lyrics <- all_lyrics_dt[mxm_id %in% song][, 2] %>%
          gsub("\\(Feat. P. Diddy\\)", "", .) %>% 
          gsub("P- ", "", .) %>%
          gsub("N- ", "", .)

# replace lyrics
all_lyrics_dt[mxm_id %in% song][, 2] <- lyrics
```

We remove artist names in the first song:

```{r export_lyrics_2.2}
primary_data_path <- here("IV_survey_builder", "primary_data_")
file_name <- paste0(primary_data_path, "/", "not_simple_medium_alpha_2021_full_2_2.csv")

write.csv(all_lyrics_dt, file_name, na="", row.names = FALSE)
```

Here we document manual changes to song lyrics based on discussions among the research group, resulting in primary data __version 2.1__. These manual changes include:
* the minor adjustment of non-English characters, 
* the removal of time-stamps, 
* the removal of artist and/or song names, and 
* the removal of section indications (e.g. 'verse', 'chorus' etc.). 

At this time we found 7 songs that could be manually adjusted for inclusion in our study, denoted by the 'A' indicator in our spreadsheet. 

```{r import_files_2.1}
# path with data
data_file_path <- here("IV_survey_builder", "primary_data_")
lyrics_file    <- "not_simple_medium_alpha_2021_full_2.csv"
selection_file <- "not_simple_medium_alpha_2021_full_10w (2).csv"

# manual selection process
selected_dt <- fread(here(data_file_path, selection_file)) %>% 
  select(mxm_id, Keep) %>% 
  # these are the four identifiers from the lyric rating process
  filter(Keep == "A" | Keep == "a")

# make smaller dt with only adjustable songs
songs <- selected_dt[,1]

# lyrics datasets
all_lyrics_dt     <- fread(here(data_file_path, lyrics_file)) %>% select(mxm_id, lyrics_body)
```

We replace non-English characters in the first song:

```{r first_song_2.1}
# select first song
song <- as.character(songs[1,1])

#view complete lyrics
knitr::kable(all_lyrics_dt[mxm_id %in% song][,2])

# replace characters
lyrics <- all_lyrics_dt[mxm_id %in% song][, 2] %>%
          gsub("抰", "'t", .) %>% 
          gsub("抯", "'s", .) %>%
          gsub("抮", "'r", .)

# replace lyrics
all_lyrics_dt[mxm_id %in% song][, 2] <- lyrics
```

We remove artist and song name from the second song:

```{r second_song_2.1}
# select second song
song <- as.character(songs[2,1])

#view complete lyrics
knitr::kable(all_lyrics_dt[mxm_id %in% song][,2])

# replace characters
lyrics <- all_lyrics_dt[mxm_id %in% song][, 2] %>%
  gsub("My silent loveisham jones ", "", .)

# replace lyrics
all_lyrics_dt[mxm_id %in% song][, 2] <- lyrics
```

We remove song section from the third song:

```{r third_song_2.1}
#select song
song <- as.character(songs[3,1])

#view complete lyrics
knitr::kable(all_lyrics_dt[mxm_id %in% song][,2])
#or copy this to console:
print(all_lyrics_dt[mxm_id %in% song][,2])

lyrics <- all_lyrics_dt[mxm_id %in% song][, 2] %>%
  gsub("\\[Verse\\]", "", .) %>%
  gsub("\\(Choppa\\)", "", .) %>%
  gsub("\\(Verse One\\)", "", .)

# replace lyrics
all_lyrics_dt[mxm_id %in% song][, 2] <- lyrics
```

We replace artist names from fourth song:

```{r fourth_song_2.1}
#select song
song <- as.character(songs[4,1])

#view complete lyrics
knitr::kable(all_lyrics_dt[mxm_id %in% song][,2])
#or copy this to console:
print(all_lyrics_dt[mxm_id %in% song][,2])

lyrics <- all_lyrics_dt[mxm_id %in% song][, 2] %>%
  gsub("Chad: ", "", .) %>%
  gsub("Chad and Team: ", "", .) %>%
  gsub("Ryan: ", "", .) %>%
  gsub("Chad and Team: ", "", .) %>%
  gsub("Girls in the Stands: ", "", .) %>%
  gsub("Ryan and Team: ", "", .) %>%
  gsub("Ryan, Team and Girls:", "", .) %>%
  gsub("Team: ", "", .)

# replace lyrics
all_lyrics_dt[mxm_id %in% song][, 2] <- lyrics
```

We replace time stamps in the fifth song:

```{r fifth_song_2.1}
#select song
song <- as.character(songs[5,1])

#view complete lyrics
knitr::kable(all_lyrics_dt[mxm_id %in% song][,2])
#or copy this to console:
print(all_lyrics_dt[mxm_id %in% song][,2])

lyrics <- all_lyrics_dt[mxm_id %in% song][, 2] %>%
  gsub("\\[00: 16.31\\]", "", .) %>%
  gsub("\\[00: 19.24\\]", "", .) %>%
  gsub("\\[00: 21.99\\]", "", .) %>%
  gsub("\\[00: 27.72\\]", "", .) %>%
  gsub("\\[00: 30.41\\]", "", .) %>%
  gsub("\\[00: 33.21\\]", "", .) %>%
  gsub("\\[00: 39.40\\]", "", .) %>%
  gsub("\\[00: 42.20\\]", "", .) %>%
  gsub("\\[00: 44.89\\]", "", .) %>%
  gsub("\\[00: 50.63\\]", "", .) %>%
  gsub("\\[00: 53.43\\]", "", .) %>%
  gsub("\\[00: 56.24\\]", "", .) %>%
  gsub("\\[01: 24.08\\]", "", .) %>%
  gsub("\\[01: 27.00\\]", "", .) %>%
  gsub("\\[01: 29.57\\]", "", .)

# replace lyrics
all_lyrics_dt[mxm_id %in% song][, 2] <- lyrics
```

We remove artist names from the sixth song:

```{r sixth_song_2.1}
#select song
song <- as.character(songs[6,1])

#view complete lyrics
knitr::kable(all_lyrics_dt[mxm_id %in% song][,2])
#or copy this to console:
print(all_lyrics_dt[mxm_id %in% song][,2])

lyrics <- all_lyrics_dt[mxm_id %in% song][, 2] %>%
  gsub("Featuring Joe Stampley", "", .)

# replace lyrics
all_lyrics_dt[mxm_id %in% song][, 2] <- lyrics
```

We remove artist names from the seventh song:

```{r seventh_song_2.1}
#select song
song <- as.character(songs[7,1])

#view complete lyrics
knitr::kable(all_lyrics_dt[mxm_id %in% song][,2])
#or copy this to console:
print(all_lyrics_dt[mxm_id %in% song][,2])

lyrics <- all_lyrics_dt[mxm_id %in% song][, 2] %>%
  gsub("\\(Feat. Method Man\\)", "", .) %>%
  gsub("\\[N-Tyce\\])", "", .) %>%
  gsub("\\[Chorus: N-Tyce \\(Method Man\\)\\]", "", .)

# replace lyrics
all_lyrics_dt[mxm_id %in% song][, 2] <- lyrics
```

We write primary data file version 2.1, which includes the manual adjustments. 

```{r export_lyrics_2.1}
primary_data_path <- here("IV_survey_builder", "primary_data_")
file_name <- paste0(primary_data_path, "/", "not_simple_medium_alpha_2021_full_2_1.csv")

write.csv(all_lyrics_dt, file_name, na="", row.names = FALSE)
```


