---
title: "prolific integration test"
author: "Andrew M. Demetriou"
date: "12/21/2021"
---

```{r setup}
library("here")
library("data.table")
library("dplyr")
```

#AIM:

To check whether we've correctly coded the spreadsheet to automatically collect participant identification numbers from prolific.co, we put together a pilot. 

The pilot will ask participants to manually enter their prolific ID, which we will check against the automatically collected prolific ID. We include all the components of our survey at the time, but only included a small number of survey items. 

```{r create_header_footer}
# path with data
data_file_path <- here("template_files_")

#complete template for survey
complete_survey_dt <- fread(here(
  data_file_path, "lyrics_annotation_template_v9 - survey.csv"))

# set survey 'header' and 'footer', without lyric annotation components
header_dt <- complete_survey_dt %>% .[1:29]
footer_dt <- complete_survey_dt %>% .[51]

# bind header and footer
survey_dt <- rbind(header_dt, footer_dt)
```

We upload items to formR via csv. This chunk saves the .csv to export for testing on prolific. 

```{r export survey}
# folder to save data
data_file_path <- here("IV_survey_builder", "survey_files_")

# adjust file name based to save based on convention and number
file_name <- paste0(data_file_path, "/", "prolific_integration_test_v2.csv")
  
# write file
#write.csv(survey_dt, file_name, na="", row.names = FALSE)
```

