---
title: "Demographics"
author: "andrew demetriou"
date: "26/04/2023"
---

#AIM

The aim of this notebook is to explore the demographic variables provided by prolific. 
The dataframe used here is not in the public repo. It includes both complete and partial responses for which individuals indicated consent on the consent form. 

```{r setup, include=FALSE}
library('here')       # file logistics
library('data.table') # data manipulation
library('dplyr')      # data manipulation
library('ggplot2')    # data visualization
library('gridExtra')  # data visualization
library('naniar')     # visualizing missingness
library('forcats')    # ordering factors

options(scipen=999)
```

```{r}
# load dataset
load(here("VI_data_collection", "_data", "survey_response_data", "working_df.RDS"))

# create a working dataset
responses_dt <- as.data.table(working_df)

rm(working_df)
```

```{r}
# examine missingess of variables
missingness_dt <- responses_dt[,participant_ID:Finished] |> unique()

# make a summary table
missingness_dt |> 
  group_by(Finished) |>
  select(-starts_with("Lyric")) |>
  miss_var_summary()

# demographic data for surveys that were not completed is mostly unavailable

rm(missingness_dt)
```

Less than 20% of respondents left feedback. Respondents reside in the US according to Prolific. 


```{r}
# coerce to numeric
responses_dt$Age <- as.numeric(responses_dt$Age)

age_dt <- responses_dt |> 
  select(participant_ID, Age) |>
  base::unique()

# median age is 40
median(age_dt$Age, na.rm=TRUE)

# plot histogram
age_dt |>
  ggplot(aes(x=Age)) +
  geom_histogram(bins=50) +
  geom_vline(xintercept = 40, color="red") +
  geom_vline(xintercept = 18, color="blue") +
  geom_vline(xintercept = 77, color="blue") +
  annotate(geom="text", x=46, y=30, label="Median = 40") +
  annotate(geom="text", x=22, y=30, label="Min = 18") +
  annotate(geom="text", x=72, y=30, label="Max = 77") +
  xlab("Respondent Age") +
  ylab("") +
  theme_minimal()

rm(age_dt)
```
```{r}
# build demogs dt
demogs_dt <- responses_dt %>%
  select(participant_ID, Age:Finished) |>
  unique()
```

```{r}
# function to plot various demographic columns
plot <- function(column){
  
  # use this to convert string to an accessible object
  column <- sym(column)
  
  demogs_dt %>%
    group_by(!!column) %>%
    # create a column that counts the different responses
    summarize(counts = n()) %>%
      ggplot(aes(x=!!column, fill = !!column, y=counts)) +
      geom_bar(stat = 'identity') +
      # add a text entry to show the count on the plot
      geom_text(aes(label = counts)) + 
      coord_flip() +
      theme_minimal() +
      scale_x_discrete(position = "top") +
      xlab("") +
      labs(title = column) +
      theme(legend.position = "none")
}
```

There's a lot of missingness in this dataset in terms of the demographics:


```{r}
plot("Employment status")
```

Still, it seems clear most identify as 'white'. 

```{r}
plot("Ethnicity simplified")
```

Most were born in the US. 

```{r}
plot("Country of birth")
```

This seems to be the variable that determines the sample for prolific; responses are either US or NA. 

```{r}
plot("Country of residence")
```

Most identify as US American. 

```{r}
plot("Nationality")
```

Most speak English as their first language. 

```{r}
plot("Language")
```

Of those whose data is available, the Sex appears to be split. 

```{r}
plot("Sex")
```

