---
title: "estimate intercepts using lme4 and mplus"
author: "andrew demetriou"
---

```{r setup, include=FALSE}
library('here')            # file logistics
library('data.table')      # data manipulation
library('dplyr')           # data manipulation
library("MplusAutomation") # Mplus integration

# load and format datasets
source(here("VI_data_collection", "3_response_intercept_analysis", "0_format_response_data_1.1.R"))
```

```{r}
# dataframe with all responses
responses_dt_a <- rbind(
  responses_dt   %>% select(participant_ID, item_ID, all_of(values)), 
  responses_dt_p %>% select(participant_ID, item_ID, all_of(values)))

responses_dt_a$row_number <- seq(1:nrow(responses_dt_a))
```

We have three datasets loaded:

From current survey:
*`responses_dt_c` is only complete responses from the current survey
*`responses_dt` is both all complete and all partial responses from current survey
*`responses_dt_p` is all complete responses from the pilot study
*`responses_dt_a` is all responses from current survey and pilot study

```{r}
format_df_for_mplus <- function(df){
  # select only ID columns and values scores
  mplus_df <- df %>% select(row_number, participant_ID, item_ID, all_of(values))
  
  # rename columns for Mplus
  colnames(mplus_df) <- c("row_number", "subject_ID", "item_ID", paste0("y", seq(1, 10)))
  
  # coerce to dataframe for mplus modeller
  mplus_df <- as.data.frame(mplus_df)
  
  return(mplus_df)
}
```

```{r}
fit_mplus_model <- function(df){
  
  # run mplus model
 fit <- mplusModeler(mplus_model, 
                     modelout = "VI_data_collection/_data/mplus_output/participant_model_1.0.inp", 
                     hashfilename = FALSE,
                     check = TRUE,
                     run=TRUE)

  # print errors and warnings
  print(fit$results$errors)
  print(fit$results$warnings)
  
  return(fit)
}
```

This wouldn't run in a single function, so I run it iteratively with each dataset:

```{r}
# format df
mplus_df <- format_df_for_mplus(responses_dt)  
  
# load mplus model specification
source(here("VI_data_collection", "2_intermediary_response_data_computing",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit <- mplus_df %>% fit_mplus_model()

#saveRDS(fit_responses_dt, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_1.0.RDS"))
```

```{r}
# format df
mplus_df <- format_df_for_mplus(responses_dt_p)  
  
# load mplus model specification
source(here("VI_data_collection", "2_intermediary_response_data_computing",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_p <- mplus_df %>% fit_mplus_model()

#saveRDS(fit_responses_dt, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_p_1.0.RDS"))
```

```{r}
# format df
mplus_df <- format_df_for_mplus(responses_dt_c)  
  
# load mplus model specification
source(here("VI_data_collection", "2_intermediary_response_data_computing",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_c <- mplus_df %>% fit_mplus_model()

#saveRDS(fit_responses_dt, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_c_1.0.RDS"))
```
```{r}
# format df
mplus_df <- format_df_for_mplus(responses_dt_a)  
  
# load mplus model specification
source(here("VI_data_collection", "2_intermediary_response_data_computing",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_a <- mplus_df %>% fit_mplus_model()

#saveRDS(fit_responses_dt_a, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_a_1.0.RDS"))
```
```{r}
# load mplus fit objects saved from notebook 1
fit   <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_1.0.RDS"))
fit_a <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_a_1.0.RDS"))
fit_c <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_c_1.0.RDS"))
fit_p <- readRDS(here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_p_1.0.RDS"))
```


```{r}
# load mplus functions
source(here("VI_data_collection", "2_intermediary_response_data_computing", "_functions", "mplus_descriptives_function_1.7.R"))

# load lme4 functions
source(here("VI_data_collection", "2_intermediary_response_data_computing", "_functions", "lme4_descriptives_function_1.7.R"))
```

```{r}
build_intercept_df <- function(df, fit_object=fit, dataset_label){
  
  # df <- responses_dt_p %>% rename(subject_ID = participant_ID)
  # fit_object <- fit_p
  
  df <- get_random_intercepts_mplus(
    # adjust column name to fit legacy function
    df %>% rename(subject_ID = participant_ID), fit_object, 
    # specify intercept to retrieve
    intercept = 'item_ID') %>%
    # pivot values columns to long format
    pivot_longer(cols = all_of(values), names_to = "value", values_to = "intercept") %>% 
    # reorder df
    arrange(value, ITEM_ID) %>%
    # add column that labels the dataset
    mutate(dataset = dataset_label)
}
```

```{r}
# build intercept df for each dataset
mplus_intercept_df_1 <- build_intercept_df(responses_dt, fit, "complete and partial responses") %>% rename(item_ID = ITEM_ID)
mplus_intercept_df_c <- build_intercept_df(responses_dt_c, fit_c, "complete responses only") %>% rename(item_ID = ITEM_ID)
mplus_intercept_df_p <- build_intercept_df(responses_dt_p, fit_p, "pilot study responses only") %>% rename(item_ID = ITEM_ID)
mplus_intercept_df_a <- build_intercept_df(responses_dt_a, fit_a, "all responses") %>% rename(item_ID = ITEM_ID)
```

```{r}
lme4_intercept_and_predicted <- function(df){
  # select only ID columns and values scores
  df <- df %>% select(participant_ID, item_ID, all_of(values))
  
  # rename columns to have same names as in functions
  colnames(df) <- c("subject_ID", "item_ID", values)
  
  # run lmer models with each value as the dependent variable and
  # extract standard deviation 95% CI for participants and songs by value
  lme4_intercept_and_predicted_dt <- df |> 
     assemble_intercept_and_predicted_dt()
}
```

```{r}
lme4_intercept_dt   <- lme4_intercept_and_predicted(responses_dt)   %>% 
  mutate(dataset = "complete and partial responses")
lme4_intercept_dt_c <- lme4_intercept_and_predicted(responses_dt_c) %>% 
  mutate(dataset = "complete responses only")
lme4_intercept_dt_p <- lme4_intercept_and_predicted(responses_dt_p) %>% 
  mutate(dataset = "pilot study responses only")
lme4_intercept_dt_a <- lme4_intercept_and_predicted(responses_dt_a) %>% 
  mutate(dataset = "all responses")
```

```{r}
df <- merge(lme4_intercept_dt, mplus_intercept_df_1, by = c("item_ID", "value"))
cor(df$intercept.x, df$intercept.y)

df <- merge(lme4_intercept_dt_c, mplus_intercept_df_c, by = c("item_ID", "value"))
cor(df$intercept.x, df$intercept.y)

df <- merge(lme4_intercept_dt_p, mplus_intercept_df_p, by = c("item_ID", "value"))
cor(df$intercept.x, df$intercept.y)

df <- merge(lme4_intercept_dt_a, mplus_intercept_df_a, by = c("item_ID", "value"))
cor(df$intercept.x, df$intercept.y)
```


