---
title: "run mplus participant model"
author: "andrew demetriou"
---

```{r setup, include=FALSE}
library('here')            # file logistics
library('data.table')      # data manipulation
library('dplyr')           # data manipulation
library("MplusAutomation") # Mplus integration

# load and format datasets
source(here("VI_data_collection", "2_intermediary_response_data_computing", "0_format_response_data_1.1.R"))
```

```{r}
# dataframe with all responses
responses_dt_a <- rbind(
  responses_dt   %>% select(participant_ID, item_ID, all_of(values)), 
  responses_dt_p %>% select(participant_ID, item_ID, all_of(values)))

responses_dt_a$row_number <- seq(1:nrow(responses_dt_a))
```

We have three datasets loaded:

From current survey:
*`responses_dt_c` is only complete responses from the current survey
*`responses_dt` is both all complete and all partial responses from current survey
*`responses_dt_p` is all complete responses from the pilot study
*`responses_dt_a` is all responses from current survey and pilot study

```{r}
format_df_for_mplus <- function(df){
  # select only ID columns and values scores
  mplus_df <- df %>% select(row_number, participant_ID, item_ID, all_of(values))
  
  # rename columns for Mplus
  colnames(mplus_df) <- c("row_number", "subject_ID", "item_ID", paste0("y", seq(1, 10)))
  
  # coerce to dataframe for mplus modeller
  mplus_df <- as.data.frame(mplus_df)
  
  return(mplus_df)
}
```

```{r}
fit_mplus_model <- function(df){
  
  # run mplus model
 fit <- mplusModeler(mplus_model, 
                     modelout = "VI_data_collection/_data/mplus_output/participant_model_1.0.inp", 
                     hashfilename = FALSE,
                     check = TRUE,
                     run=TRUE)

  # print errors and warnings
  print(fit$results$errors)
  print(fit$results$warnings)
  
  return(fit)
}
```

This wouldn't run in a single function, so I run it iteratively with each dataset:

```{r}
# format df
mplus_df <- format_df_for_mplus(responses_dt)  
  
# load mplus model specification
source(here("VI_data_collection", "2_intermediary_response_data_computing",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit <- mplus_df %>% fit_mplus_model()

saveRDS(fit, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_1.0.RDS"))
```

```{r}
# format df
mplus_df <- format_df_for_mplus(responses_dt_p)  
  
# load mplus model specification
source(here("VI_data_collection", "2_intermediary_response_data_computing",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_p <- mplus_df %>% fit_mplus_model()

saveRDS(fit_p, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_p_1.0.RDS"))
```

```{r}
# format df
mplus_df <- format_df_for_mplus(responses_dt_c)  
  
# load mplus model specification
source(here("VI_data_collection", "2_intermediary_response_data_computing",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_c <- mplus_df %>% fit_mplus_model()

saveRDS(fit_c, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_c_1.0.RDS"))
```
```{r}
# format df
mplus_df <- format_df_for_mplus(responses_dt_a)  
  
# load mplus model specification
source(here("VI_data_collection", "2_intermediary_response_data_computing",
            "_mplus_models", "participant_model_1.0.R"))

# fit model
fit_a <- mplus_df %>% fit_mplus_model()

saveRDS(fit_a, here("VI_data_collection", "_data", "intermediary_data", "mplus_participant_model_a_1.0.RDS"))
```