---
title: "alpha_estimation"
author: "andrew demetriou"
date: "07/10/2022"
---

# AIM:

The aim of this notebook is to estimate the number of ratings necessary per item. The threshold is estimated using cronbach's alpha. 

```{r setup, include=FALSE}
library('data.table') # data manipulation
library('here')       # file logistics
library('dplyr')      # logistics
library('ggplot2')    # visualization
library('shiny')      # interactive results
library('DT')         # interactive table
```

Import dataset in qualtrics format, and re-organize for further processing:

```{r}
# organize qualtrics output
source(here("II_rater_pilot", "scripts", "data_file_re-shape_1.0.R"))

# path with the actual data
data_file_path <- here("II_rater_pilot", "response_data")

# file name as a string
file_name <- "annotation_number_estimation _2.2_November 28, 2021_07.51.csv"

# read in qualtrics data file
responses_dt <- fread(here(data_file_path, file_name)) %>%
  data_file_reshape(., file_name)

rm(data_file_path, file_name, data_file_reshape)
```

The general approach is to compute the cronbach's alpha per value. Participants responded on a scale of -100 to 100. Because 0 is a legitimate response, issues may arise with further computations. Thus, I add 101 to the scores. 

```{r}
responses_dt[,3:12] <- responses_dt[,3:12] + 101
```

```{r}
# import custom alpha estimation function
source(here("0_functions", "alpha_function_1.9.R"))
```

```{r}
# make a dataframe of alpha estimations for each variable
list_of_alpha_tibbles <- make_list_of_alpha_tibbles(responses_dt)

# summarize dataframes
alpha_tibble_summary <- make_tibble_of_alpha_tibble_summaries(responses_dt)

# smallest n that reaches threshold
minimum_n <- minimum_n_by_alpha(responses_dt)

rm(get_alpha, make_list_of_alpha_tibbles, pivot_dataframe, make_list_of_pivoted_dataframes, rename_columns, samples_of_alphas, summarize_alpha_tibble, make_tibble_of_alpha_tibble_summaries, minimum_n_by_alpha)
```


```{r}
plot_alphas <- function(alpha_dt){
  ggplot(alpha_dt) +
    geom_vline(xintercept = 0.7, color = "grey50" ) +
    geom_density(aes(alpha, colour = as.factor(n))) + 
    xlim(0, 1) +
    labs(title = "Distribution of Alphas by Rater N\n", 
       color = "Number of raters\n") +
    theme_minimal()
}
```

```{r}
list_of_plots <- lapply(list_of_alpha_tibbles, plot_alphas)
names <- names(list_of_alpha_tibbles)
```

```{r}
for(i in 1:length(names)){
  list_of_plots[[i]] <- list_of_plots[[i]] + labs(title=paste("Distribution of Alphas by Rater N for", names[i]))
}
```

```{r}
list_of_plots |>
  saveRDS(here("II_rater_pilot", "intermediary_data", "list_of_alpha_plots.RDS"))

list_of_alpha_tibbles |>
  saveRDS(here("II_rater_pilot", "intermediary_data", "list_of_alpha_tibbles.RDS"))
```


```{r}
ui <- fluidPage(
  titlePanel("Alpha Reliability by Value"),
  
  # drop down menu to display the values
  selectInput('value', 'select value', choice = c(names)), 
  
  #plot this thing called 'alpha_plot'
  plotOutput('alpha_plot'),
  
  #tableOutput('alpha_table')
  DT::DTOutput('alpha_table')
)

server <- function(input, output, session){
  
  # render the plots; alpha distribution by n
  output$alpha_plot <- renderPlot({
  
    #note: I've already created the plots 
    plot_alphas(list_of_alpha_tibbles[[input$value]]) +
    
    #this adds the label to the title
    labs(title=paste("Distribution of Alphas by Rater N for", input$value))
  })
  
  # render the table of minimum n alpha summaries
  output$alpha_table <- renderDataTable(minimum_n)
}
shinyApp(ui=ui, server=server)
```

